Formalizing Stack Safety as a Security Property

### 2021-07-08 submission3/main.tex
Section 3 work

@@

\section{Machines, Enforcement, and Threat Model}
\label{sec:setup}

Before introducing the model\apt{what does this really mean?}  itself, we set the context by stating some
assumptions we are making about machines, enforcement mechanisms, and
potential threats.

\paragraph*{Machine model}
Our properties are defined in terms of an abstract machine model that makes very few
assumptions about the concrete machine that instantiates it. Without changing the
properties, we can support\apt{what does this really mean?} any register
machine with a program counter (\(\PCname\)) and a stack pointer (\(\rsp\)); other
details of the machine's ISA are unimportant. We can support any static memory,
including a heap---our properties define stack protection regardless of how
non-stack memory is used. Likewise, we can support a range of calling convention
choices, like whether or not to have a frame pointer.

Some calling convention choices, such as callee-saves registers and argument
passing on the stack, require natural enhancements to our base properties.\apt{
  At this point it is unclear what these ``enhancements'' might be.}
We
present these enhancements, as well as a simple coroutine model,
in Section~\ref{sec:ext}, after we formalize the essence of stack safety.
Other plausible enhancements include:\apt{Better to move the ``what would
  be needed'' parts to the future work section -- they are mysterious here.}

\begin{itemize}
\item address-taken locals---similar to argument passing, but
      would benefit from more sophisticated, heap-like memory models
\item exceptions---needs annotations of throw/catch points
\item continuations---would be similar to argument passing
\end{itemize}

Runtime code generation would be much more challenging to support, as
it requires dynamic generation of code annotations. Similarly, true concurrency
would require more substantial changes to the machine
model. We consider these features out of scope for now.

@@

Dividing policy state from machine state helps us reason about
enforcement mechanisms that keep a significant amount of data separate from the
primary execution data to inform security. Obviously tag-based systems do this\apt{At this
  point, we barely know what tags are},
with tags and extra policy state separated from, and untouchable by, application code.
Another significant example is the shadow stack\apt{citations? what enforcement systems do we actually have in mind?}, where a second stack
mirrors the real one, and a return may only occur if both stacks agree on the
target address. Typically the shadow stack is isolated using a separate protection
mechanism.

The separation is smaller in protection mechanisms that interact more directly
with application code.\apt{Vague. Mention capabilities?}  Some mechanisms might exhibit failstop behavior, but have
no protected policy state; others might never even failstop. Then the policy-enhanced
machine is simply identical to the original. This is the case for software-based
techniques.\apt{Even if they use a shadow stack?}

\paragraph*{Threat Model}
\label{sec:threat}

Consider a common scenario: a developer wants to link against an existing
code-base, which will then occupy the same address space as their own code.
But perhaps the external code is buggy, or even malicious, and contains
exploits that can be used to cause it to modify the developer's data or
exfiltrate secrets. For that matter, the developer may not experienced at
secure programming and so may not be able to rely on their own skill
to keep their secrets safe.

The developer likely relies instead on simple reasoning principles about how
their code should work. Local variables do not change across function calls,
and the callees generally do the same thing regardless of when and where they
are called (up to a reasonable interface, such as their parameters and global
variables.) If these intuitions prove false, then the programmer will reason
incorrectly about their program and especially about its security, and
be exposed to attack.

Intuitively, stack safety means that the stack frame of each caller is protected
from attempts by a confused or malicious callee (including all its
descendant callees) to read or write its private data.
That is, the attacker model treats every caller as a defender and its callee as
an attacker, meaning that with nested calls, a single function invocation
may be viewed in both ways.
\apt{everything down to here seems like it should go earlier -- in informal section
  or even intro.}
We do rule out some
scenarios with a ``well-formedness'' criterion on initial states, ensuring that
they are set up properly for the enforcement mechanism in question.
When validating \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies,
well-behaved calls and returns are implemented by fixed sequences of instructions
which the tag policy will force to be executed from start to finish,
termed {\em blessed sequences}. Well-formedness requires that all annotated
instructions will be situated at the appropriate instruction of a {\em blessed sequence}.
This is all we need for the policy to protect itself: it can then failstop if it would
execute the blessed sequence out of order.

We adopt a strong threat model in which both caller and callee may execute
arbitrary machine code. \apt{This should be the lead sentence. But how does is
  it consistent with the previous para on well-formedness?}
  Hardware and timing attacks are out of scope, as are
are attacks on the heap---though we believe that the work of
\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with ours
to protect both stack and heap.

>>>>

Intuitively, stack safety means that the stack frame of each caller is protected
from attempts by a confused or malicious callee (including all its
descendant callees) to read or write its private data.
That is, the attacker model treats every caller as a defender and its callee as
an attacker, meaning that with nested calls, a single function invocation
may be viewed in both ways.

@@

\section{Threat Model, Machines, and Enforcement }
\label{sec:setup}

We will introduce a formal model of stack safety that abstracts over
machine details. But first we must discuss the bounds of our assumptions ---
how broadly we can abstract over machines, how we represent enforcement, and
our threat model.

We adopt a strong threat model in which both caller and callee may execute
arbitrary machine code. But hardware and timing attacks are out of scope, as are
are attacks on the heap---though we believe that the work of
\citet{DBLP:conf/post/AmorimHP18} can be combined smoothly with ours
to protect both stack and heap.

\paragraph*{Machine model}
Our properties are defined in terms of an abstract machine model that makes very few
assumptions about the concrete machine that instantiates it. Each formal property
may be instantiated with any machine that meets the following criteria, and it
will be a meaningful defition of stack safety for that machine. It must at least be a
register machine with a program counter (\(\PCname\)), and there must be a well-defined
mechanism for determining the bounds of a stack frame, and for recognizing when a call
has returned. In this paper we discuss those machines with a stack pointer (\(\rsp\)),
in which the current stack frame is everything below the stack pointer and above the
previous frame, and a call from instruction \(\addr_\PCname\) with stack pointer \(\addr_\rsp\)
is said to have returned when the pc is at \(\addr_\PCname + 4\) and the stack pointer
is again at \(\addr_\rsp\). But, other configurations, such as those involving frame pointers,
are equally feasible. A machine may use its non-stack memory in any way, including
as a heap, as long as the stack has static bounds.

Our core property assumes that all register values passed to the callee are public,
and that all data in a stack frame is private.
Some calling convention choices, such as callee-saves registers and argument
passing on the stack, are not supported directly by this property; we introduce
secondary properties that use the same underlying formalism to handle these cases,
as well as a simple coroutine model, in Section~\ref{sec:ext}.

Other enhancements, such as runtime code generation and true concurrency,
would be much more challenging to support, and we consider these features
out of scope for now.

@@

Dividing policy state from machine state helps us reason about
enforcement mechanisms that keep a significant amount of data separate from the
primary execution data to inform security. Tag-based systems, in which a hardware
monitor tracks metadata tags associated with a software-defined ``micro-policy,'' are
the motivating application. The tags, and any extra policy state, are separated from
and untouchable by application code.
%Another significant example is the shadow stack\apt{citations? what enforcement systems do we actually have in mind?}, where a second stack
%mirrors the real one, and a return may only occur if both stacks agree on the
%target address. Typically the shadow stack is isolated using a separate protection
%mechanism.
\sna{Commented out shadow stacks until I can find a citation. May just skip it.}

That said, our model is meant to apply generically to any enforcement mechanism,
including those that do not take advantage of this separate.
Some mechanisms, like hardware capability systems, might exhibit failstop behavior,
but have no protected policy state --- this is represented by a ``unit'' policy
that failstops when the hardware fault. Others might never even failstop, such as
code rewriting techniques that turn stack safety violations into other (erroneous, but not
stack-unsafe) behaviors. Then the policy-enhanced machine is identical to the original.
\sna{I'm thinking of SFI here, but that's not stacks, right? Does Softbound work like
  this.}

An enforcement model may additionally
apply a ``well-formedness'' criterion on initial states, ensuring that
they are set up properly for the enforcement mechanism in question.
When validating \citeauthor{DBLP:conf/sp/RoesslerD18}'s micro-policies,
well-behaved calls and returns are implemented by fixed sequences of instructions
which the tag policy will force to be executed from start to finish,
termed {\em blessed sequences}. Well-formedness requires that all annotated
instructions will be situated at the appropriate instruction of a {\em blessed sequence}.
This is all we need for the policy to protect itself: it can then failstop if it would
execute the blessed sequence out of order.
